---
title: "Leave a trace: Recursive reasoning about deceptive behavior"
author: "Verona Teo, Sarah A. Wu, Erik Brockbank, and Tobias Gerstenberg"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    highlight: tango
---

# Setup

```{r setup, include = F, message = F, warning = F}
# load relevant libraries and functions
require(knitr)         # for knitting        
library(ggstatsplot)   # extension of ggplot2
library(Metrics)       # for RMSE
library(DT)            # for datatable() function 
library(brms)          # for Bayesian regression models
library(emmeans)       # for linear contrasts
library(Hmisc)         # for bootstrapped confidence intervals
library(scales)        # hue palette
library(patchwork)     # for multiple plots
library(png)           # for working with images
library(grid)          # for grob
library(egg)           # for geom_custom
library(transport)     # for Wasserstein distance
library(MASS)          # for fitting distributions to data
library(tidyverse)     # for everything else

knitr::opts_chunk$set(echo = T, warning = F, message = F)

# set default plot theme 
theme_set(theme_classic() + 
            theme(text = element_text(size = 14)))

# suppress warnings about grouping 
options(dplyr.summarise.inform = F)
options(xtable.floating = F)
options(xtable.timestamp = "")

set.seed(1)
```

```{r include = F}
trials = paste0('snack', setdiff(1:10, 4))
agents = c('A', 'B')
agent_types = c('naive', 'sophisticated')
path_types = c("path to fridge", "return path")

data_dir = '../../data/'
figures_dir = '../../figures/'
images_dir = '../trials/'

dodge_width = 0.9
agent_type_colors = c('naive' = 'darkorange2',
                      'sophisticated' = 'deepskyblue2')
path_type_alphas = c(0.4, 0.8)
model_labels = c('human' = 'Humans',
                 'rsm' = 'RSM',
                 'emp' = 'Empirical',
                 'emp_mis' = 'Mismatched empirical',
                 'gpt4o' = 'GPT-4o',
                 'unif' = 'Uniform simulation',
                 'heu' = 'Heuristic')
```

```{r include = F}
# Issue with patchwork collect guides (see (this issue)[https://github.com/thomasp85/patchwork/issues/404])

guides_build_mod <- function (guides, theme){
  legend.spacing.y <- calc_element("legend.spacing.y", theme)  # modified by me
  legend.spacing.x <- calc_element("legend.spacing.x", theme)  # modified by me
  legend.box.margin <- calc_element("legend.box.margin", theme) %||% 
    margin()
  widths <- exec(unit.c, !!!lapply(guides, gtable_width))
  heights <- exec(unit.c, !!!lapply(guides, gtable_height))
  just <- valid.just(calc_element("legend.box.just", theme))
  xjust <- just[1]
  yjust <- just[2]
  vert <- identical(calc_element("legend.box", theme), "horizontal")
  guides <- lapply(guides, function(g) {
    editGrob(g, vp = viewport(x = xjust, y = yjust, just = c(xjust, 
                                                             yjust), height = if (vert) 
                                                               heightDetails(g)
                              else 1, width = if (!vert) 
                                widthDetails(g)
                              else 1))
  })
  guide_ind <- seq(by = 2, length.out = length(guides))
  sep_ind <- seq(2, by = 2, length.out = length(guides) - 1)
  if (vert) {
    heights <- max(heights)
    if (length(widths) != 1) {
      w <- unit(rep_len(0, length(widths) * 2 - 1), "mm")
      w[guide_ind] <- widths
      w[sep_ind] <- legend.spacing.x
      widths <- w
    }
  }
  else {
    widths <- max(widths)
    if (length(heights) != 1) {
      h <- unit(rep_len(0, length(heights) * 2 - 1), "mm")
      h[guide_ind] <- heights
      h[sep_ind] <- legend.spacing.y
      heights <- h
    }
  }
  widths <- unit.c(legend.box.margin[4], widths, legend.box.margin[2])
  heights <- unit.c(legend.box.margin[1], heights, legend.box.margin[3])
  guides <- gtable_add_grob(gtable(widths, heights, name = "guide-box"), 
                            guides, t = 1 + if (!vert) 
                              guide_ind
                            else 1, l = 1 + if (vert) 
                              guide_ind
                            else 1, name = "guides")
  gtable_add_grob(guides, element_render(theme, "legend.box.background"), 
                  t = 1, l = 1, b = -1, r = -1, z = -Inf, clip = "off", 
                  name = "legend.box.background")
}

environment(guides_build_mod) <- asNamespace('patchwork')
assignInNamespace("guides_build", guides_build_mod, ns = "patchwork")
```


# Helper functions

## Demographics
```{r include = F}
print_demographics = function(data) {
  # gender
  data %>%
    group_by(gender) %>%
    summarise(n = n()) %>%
    print()
  
  # age
  print(sprintf('age: M = %.0f, SD = %.0f',
                mean(data$age, na.rm = T),
                sd(data$age, na.rm = T)))
  
  # race
  data %>%
    group_by(race) %>%
    summarise(n = n()) %>%
    print()
  
  # ethnicity
  data %>%
    group_by(ethnicity) %>%
    summarise(n = n()) %>%
    print()

  # time taken
  print(sprintf('time: %.1f minutes (SD = %.1f)',
                mean(data$time, na.rm = T)/60,
                sd(data$time, na.rm = T)/60))
}
```

## Earth Mover's Distance (Wasserstein distance)
```{r}
compute_emd = function(path_data) {
  models = setdiff(unique(path_data$model), 'human')
  data = expand.grid(trial = trials,
                     agent = agents,
                     agent_type = agent_types,
                     model = models,
                     emd = NA)
  
  for (this_trial in trials) {
    for (this_agent in agents) {
      for (this_agent_type in agent_types) {
        d = path_data %>% 
          filter(agent == this_agent,
                 trial == this_trial,
                 path_type == 'return path') %>%
          group_by(agent_type, model, x, y) %>%
          summarise(n = n()) %>%
          ungroup(x, y) %>% 
          mutate(p = n / sum(n)) %>% 
          ungroup()
        d.human = d %>%
          filter(model == 'human', agent_type == this_agent_type)
        for (this_model in models) {
          d.model = d %>%
            filter(model == this_model, agent_type == this_agent_type)
          if (length(d.model$p) > 0) {
            this_w = wasserstein(wpp(data.matrix(d.human %>% select(x, y)), d.human$p),
                                 wpp(data.matrix(d.model %>% select(x, y)), d.model$p))
            data = data %>%
              mutate(emd = ifelse((trial == this_trial) &
                                    (agent == this_agent) &
                                    (agent_type == this_agent_type) &
                                    (model == this_model),
                                  this_w,
                                  emd)
              )
          }
        }
      }
    }
  }
  
  return(data)
}
```

## Miscellaneous
```{r include = F}
rmse2 = function(x, y, na.rm = TRUE){
  res = sqrt(mean((x - y)^2, na.rm = na.rm))
  return(res)
}
```

# Experiment 1 (Suspect)

## Data
```{r include = F}
import_suspect_data = function(dir_name) {
  data = data.frame()
  for (trial in trials) {
    data = data %>%
      rbind(read_csv(str_c(data_dir, 'exp1_suspect/', dir_name, trial, '.csv'),
                     show_col_types = F) %>%
              mutate(trial = sub('_.*', '', trial),
                     # remove spaces and switch () to []
                     path = gsub('\\(', '[', path),
                     path = gsub('\\)', ']', path),
                     path = gsub(' ', '', path)) %>%
              group_by(trial, agent, agent_type, is_return) %>%
              mutate(id = row_number()) %>%
              ungroup())
  }
  return(data)
}

import_suspect_model_data = function(w, temp) {
  dir_name = str_c('rsm/w', w, '_temp', temp, '/')
  return(import_suspect_data(dir_name))
}

combine_suspect_data = function(dataframes) {
  data = data.frame()
  for (df in names(dataframes)) {
    data = data %>%
      rbind(dataframes[[df]] %>%
              mutate(model = df))
  }
  data = data %>%
    mutate(agent_type = factor(agent_type, levels = agent_types),
           path_type = ifelse(is_return, "return path", "path to fridge"),
           path_type = factor(path_type, levels = path_types),
           path_len = str_count(path, "\\[\\d+,\\d+\\]"),
           trial = factor(trial)) %>%
    select(-is_return)
  return(data)
}

get_suspect_paths = function(suspect_data, jitter_amount = 0.1) {
  data = suspect_data %>%
    mutate(path = str_remove_all(path, "\\[\\["),
           path = str_remove_all(path, "\\]\\]")) %>%
    # Split by inner coordinates
    separate_rows(path, sep = "\\],\\[") %>%
    # Add brackets back for each coordinate
    mutate(tile = str_c("[", path, "]")) %>%
    # Extract x and y values
    mutate(x = as.numeric(str_extract(tile, "(?<=\\[)\\d+")),
           y = 15 - as.numeric(str_extract(tile, "\\d+(?=\\])")),
           xx = jitter(x, amount = jitter_amount) + 0.5,
           yy = jitter(y, amount = jitter_amount) - 0.5) %>%
    select(-path, -path_len) %>%
    group_by(model, id, trial, agent, agent_type, path_type) %>%
    mutate(progress = row_number()/n()) %>%
    ungroup()
  return(data)
}
```

### Trial images
```{r}
images.suspect = data.frame(trial = trials) %>%
  mutate(grob = map(.x = trial,
                    .f = ~ readPNG(str_c(images_dir, 'exp1_suspect/images/',
                                          .x, '_A1.png'))))
```

### Humans
```{r}
data.suspect.human = read_csv(str_c(data_dir, 'exp1_suspect/humans/human_trials.csv'),
                              show_col_types = F)

data.suspect.feedback = read_csv(str_c(data_dir, 'exp1_suspect/humans/human_feedback.csv'),
                              show_col_types = F)

data.suspect.feedback %>%
  select(id, agent_type, prediction_factors) %>%
  datatable(rownames = F)
```
Demographics:
```{r}
print_demographics(data.suspect.feedback)
```

### RSM

Grid search:
```{r echo = F}
w_values = seq(0.1, 0.8, 0.1)
temp_values = c(0.01, 0.03, 0.05, 0.1, 0.15, 0.2)
gs.suspect = expand.grid(w = w_values,
                         temp = temp_values,
                         emd_n = NA,
                         emd_s = NA)

for (this_w in w_values) {
  for (this_temp in temp_values) {
    if (!dir.exists(str_c(data_dir, 'exp1_suspect/rsm/w',
                          this_w, '_temp', this_temp))) {
      gs.suspect = gs.suspect %>%
        filter(!(w == this_w & temp == this_temp))
      next
    }
    d = combine_suspect_data(list(
      'human' = data.suspect.human,
      'rsm' = import_suspect_model_data(this_w, this_temp)
    ))
    d.paths = get_suspect_paths(d, jitter = 0.1)
    d.emd = compute_emd(d.paths) %>%
      group_by(agent_type) %>%
      summarise(mean = mean(emd)) %>%
      ungroup()
    gs.suspect = gs.suspect %>%
      mutate(emd_n = ifelse(w == this_w & temp == this_temp,
                            d.emd[d.emd$agent_type == 'naive', 'mean', drop = T],
                            emd_n),
             emd_s = ifelse(w == this_w & temp == this_temp,
                            d.emd[d.emd$agent_type == 'sophisticated', 'mean', drop = T],
                            emd_s))
  }
}

gs.suspect %>%
  arrange(w, temp) %>%
  datatable(rownames = F)
```

```{r}
best_w.suspect.naive = 0.7
best_temp.suspect.naive = 0.01
best_w.suspect.sophisticated = 0.7
best_temp.suspect.sophisticated = 0.05

data.suspect.rsm = import_suspect_model_data(best_w.suspect.naive,
                                             best_temp.suspect.naive) %>%
  filter(agent_type == 'naive') %>%
  rbind(import_suspect_model_data(best_w.suspect.sophisticated,
                                  best_temp.suspect.sophisticated) %>%
          filter(agent_type == 'sophisticated'))
```

### Uniform simulation
```{r}
data.suspect.unif = import_suspect_data('uniform/') %>%
  mutate(agent_type = 'naive') %>%
  rbind(import_suspect_data('uniform/') %>%
          mutate(agent_type = 'sophisticated'))
```

### GPT-4o

```{r}
data.suspect.gpt = read_csv(str_c(data_dir, 'exp1_suspect/gpt4o/gpt4o.csv'),
                            show_col_types = F)
data.suspect.gpt %>%
  summarise(n = n(),
            n_response = sum(response_valid),
            p_response = n_response/n,
            n_path = sum(path_valid),
            p_path = n_path/n)

data.suspect.gpt = data.suspect.gpt %>%
  filter(path_valid) %>%
  mutate(is_return = (path_type == 2),
         path = gsub(" ", "", path)) %>%
  select(id=query, trial, agent_type, agent, path, is_return)
```

### Combined

```{r}
data.suspect = combine_suspect_data(list(
  'human' = data.suspect.human,
  'unif' = data.suspect.unif,
  'rsm' = data.suspect.rsm,
  'gpt4o' = data.suspect.gpt
))
data.suspect.paths = get_suspect_paths(data.suspect, 0.1)
data.suspect.emd = compute_emd(data.suspect.paths)
```

## Analysis

### Path length
```{r}
fit.suspect.path_len.human = brm(
  formula = path_len ~ 1 + agent_type + (1 | id) + (1 | trial),
  data = data.suspect %>%
    filter(model == 'human'),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.suspect.path_len.human'
)

summary(fit.suspect.path_len.human)
```

```{r}
fit.suspect.path_len.rsm = brm(
  formula = path_len ~ 1 + agent_type + (1 | trial),
  data = data.suspect %>%
    filter(model == 'sim'),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.suspect.path_len.rsm'
)

summary(fit.suspect.path_len.rsm)
```

```{r}
fit.suspect.path_len.gpt = brm(
  formula = path_len ~ 1 + agent_type + (1 | trial),
  data = data.suspect %>%
    filter(model == 'gpt4o'),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.suspect.path_len.gpt4o'
)

summary(fit.suspect.path_len.gpt)
```

### EMD comparison
```{r}
data.suspect.emd %>%
  group_by(model, agent_type) %>%
  summarise(mean = mean(emd, na.rm = T),
            low = smean.cl.boot(emd)[2],
            high = smean.cl.boot(emd)[3])

fit.suspect.emd = brm(
  formula = emd ~ 1 + agent_type + model,
  data = data.suspect.emd,
  iter = 2000,
  seed = 1,
  file = 'cache/fit.suspect.emd'
)

summary(fit.suspect.emd)
```

```{r}
fit.suspect.emd %>%
  emmeans(pairwise ~ model)
```

## Figures

### Path length
```{r fig.width = 5, fig.height = 2}
ggplot(data = data.suspect %>%
         mutate(model = factor(model,
                               levels = names(model_labels))) %>%
         group_by(model, trial, agent_type, path_type) %>%
         summarise(path_len = mean(path_len)),
       mapping = aes(x = agent_type,
                     y = path_len,
                     fill = agent_type,
                     group = path_type,
                     alpha = path_type)) +
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               linewidth = 0.1,
               position = position_dodge(width=dodge_width)) +
  geom_point(size = 2,
             shape = 21,
             color = 'gray40',
             alpha = 0.3,
             position = position_jitterdodge(dodge.width=dodge_width),
             show.legend = F) +
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               alpha = 1,
               linewidth = 1,
               position = position_dodge(width=dodge_width),
               show.legend = F) +
  scale_alpha_manual(values = path_type_alphas) +
  scale_y_continuous(name = 'Path length',
                     breaks = seq(0, 30, 10),
                     expand = expansion(mult = 0.01)) +
  scale_fill_manual(values = agent_type_colors) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 33)) +
  facet_wrap(~ model,
             nrow = 1,
             strip.position = 'bottom',
             labeller = as_labeller(model_labels)) +
  theme(text = element_text(size = 12),
        legend.position = c(0.5, 0.97),
        legend.box = 'horizontal',
        legend.direction = 'horizontal',
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.key.size = unit(8, "pt"),
        legend.margin = margin(0),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(t = 0.1, r = 0, l = 0.1, b = -0.1, unit = 'cm'),
        strip.text = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major.y = element_line())

# ggsave(str_c(figures_dir, 'suspect_path_len.pdf'), width = 5, height = 2)
```

### Path traces
```{r fig.width = 13, fig.height = 8}
path_map = function(data, images_data, title, kitchen_access) {
  g = ggplot(data = data,
         mapping = aes(x = xx,
                       y = yy,
                       color = agent_type,
                       group = id)) +
    geom_custom(data = images_data,
                mapping = aes(data = grob,
                              x = -Inf, y = -Inf,
                              color = NULL, group = NULL),
                grob_fun = function(x) rasterGrob(x,
                                                  interpolate = T,
                                                  hjust = 0,
                                                  vjust = 0)) +
    annotate('rect',
             xmin = kitchen_access[1], xmax = kitchen_access[1] + 1,
             ymin = kitchen_access[2] - 1, ymax = kitchen_access[2],
             fill = 'green',
             alpha = 0.5) +
    geom_path(alpha = 0.2, linewidth = 0.2) +
    geom_point(alpha = 0.2, size = 0.5) +
    coord_fixed() +
    facet_wrap(~ agent_type, nrow = 1) +
    scale_x_continuous(limits = c(0, 15), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 16), expand = c(0, 0)) +
    scale_color_manual(values = agent_type_colors) +
    ggtitle(title) +
    guides(color = guide_legend(override.aes = list(size = 2,
                                                    linewidth = 1,
                                                    alpha = 1))) +
    theme(plot.title = element_text(size = 16, hjust = 0.5),
          text = element_text(size = 16),
          legend.direction = 'horizontal',
          legend.title = element_blank(),
          legend.text = element_text(size = 14),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.line = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank(),
          plot.margin = margin(t = 0.1, r = 0, b = 0.1, l = 0, unit = 'cm'))
  
  return(g)
}
```

```{r fig.width = 14, fig.height = 2.5}
select_trial = 'snack2'
select_agent = 'A'
kitchen_access = c(5, 12)
path_map.human = path_map(
  data.suspect.paths %>%
    filter(trial == select_trial, model == 'human',
           agent == select_agent, path_type == 'return path'),
  images.suspect %>%
   filter(trial == select_trial),
  'Humans',
  kitchen_access
)
path_map.sim = path_map(
  data.suspect.paths %>%
    filter(trial == select_trial, model == 'rsm',
           agent == select_agent, path_type == 'return path'),
  images.suspect %>%
    filter(trial == select_trial),
  'RSM',
  kitchen_access
)
path_map.gpt = path_map(
  data.suspect.paths %>%
    filter(trial == select_trial, model == 'gpt4o',
           agent == select_agent, path_type == 'return path'),
  images.suspect %>%
    filter(trial == select_trial),
  'GPT-4o',
  kitchen_access
)
path_map.unif = path_map(
  data.suspect.paths %>%
    filter(trial == select_trial, model == 'unif',
           agent == select_agent, path_type == 'return path'),
  images.suspect %>%
    filter(trial == select_trial),
  'Uniform simulation',
  kitchen_access
)

path_map.human + path_map.sim + path_map.gpt + path_map.unif +
  plot_layout(guides = 'collect', nrow = 1) &
  theme(legend.position = 'bottom',
        legend.margin = margin(t = -0.5, b = 0, unit = 'cm'),
        )

# ggsave(str_c(figures_dir, 'suspect_paths.pdf'), width = 14, height = 2.5)
```

### EMD
```{r fig.width = 5, fig.height = 2.25}
ggplot(data.suspect.emd %>%
         mutate(model = factor(model, levels = names(model_labels))),
       aes(x = model,
           y = emd,
           fill = agent_type)) +
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               linewidth = 0.1,
               alpha = 0.8,
               position = position_dodge(width=dodge_width)) +
  geom_point(size = 2,
             shape = 21,
             color = 'gray40',
             alpha = 0.3,
             position = position_jitterdodge(jitter.width = 0.2,
                                             dodge.width = dodge_width),
             show.legend = F) +
  stat_summary(fun.data = mean_cl_boot,
               geom = "linerange",
               color = "black",
               alpha = 1,
               linewidth = 1,
               position = position_dodge(width=dodge_width),
               show.legend = F) +
  scale_fill_manual(values = agent_type_colors) +
  scale_y_continuous(name = "Earth Mover's Distance",
                     breaks = seq(0, 3, 0.5),
                     expand = expansion(mult = 0.01)) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 3.4)) +
  scale_x_discrete(labels = model_labels) +
  theme(text = element_text(size = 12),
        legend.position = c(0.5, 0.97),
        legend.direction = 'horizontal',
        legend.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.key.size = unit(8, "pt"),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(t = 0.1, r = -0.2, l = 0.1, b = 0.1, unit = 'cm'),
        panel.grid.major.y = element_line())

# ggsave(str_c(figures_dir, 'suspect_emd.pdf'), width = 5, height = 2.25)
```

# Experiment 2 (Detective)

## Data
```{r include = F}
import_detective_model_data = function(w, temp) {
  data = read_csv(str_c(data_dir, 'exp2_detective/rsm/w',
                        w, '_temp', temp, '.csv'),
                  show_col_types = F) %>%
    mutate(id = NA)
  return(data)
}

combine_detective_data = function(dataframes) {
  data = data.frame()
  for (df in names(dataframes)) {
    data = data %>%
      rbind(dataframes[[df]] %>%
              select(id, trial, agent_type, response) %>%
              mutate(model = df))
  }
  data = data %>%
    mutate(agent_type = factor(agent_type, levels = agent_types),
           trial = factor(trial))
  return(data)
}

compute_detective_means = function(detective_data) {
  data = detective_data %>%
    group_by(model, trial, agent_type) %>%
    summarise(mean = mean(response),
              low = smean.cl.boot(response)[2],
              high = smean.cl.boot(response)[3]) %>%
    ungroup() %>%
    pivot_wider(names_from = 'model',
               values_from = c('mean', 'low', 'high'))
  return(data)
}
```

### Human
```{r}
data.detective.human = read_csv(str_c(data_dir, 'exp2_detective/humans/human_trials.csv'),
                                show_col_types = F)

data.detective.feedback = read_csv(str_c(data_dir, 'exp2_detective/humans/human_feedback.csv'),
                                show_col_types = F)

data.detective.feedback %>%
  select(id, agent_type, prediction_factors) %>%
  datatable(rownames = F)
```

Demographics:
```{r}
print_demographics(data.detective.feedback)
```

### RSM

Grid search:
```{r echo = F}
w_values = seq(0.1, 0.8, 0.1)
temp_values = c(0.01, 0.03, 0.05, 0.08, 0.1, 0.12, 0.15, 0.2)
gs.detective = expand.grid(w = w_values,
                           temp = temp_values,
                           r_n = NA,
                           rmse_n = NA,
                           r_s = NA,
                           rmse_s = NA)

for (this_w in w_values) {
  for (this_temp in temp_values) {
    if (!file.exists(str_c(data_dir, 'exp2_detective/rsm/w',
                           this_w, '_temp', this_temp, '.csv'))) {
      gs.detective = gs.detective %>%
        filter(!(w == this_w & temp == this_temp))
      next
    }
    d = combine_detective_data(list(
      'human' = data.detective.human,
      'rsm' = import_detective_model_data(this_w, this_temp)
    ))
    d.means = compute_detective_means(d)
    d.means.n = d.means %>% filter(agent_type == 'naive')
    d.means.s = d.means %>% filter(agent_type == 'sophisticated')
    gs.detective = gs.detective %>%
      mutate(r_n = ifelse(w == this_w & temp == this_temp,
                          cor(d.means.n$mean_human, d.means.n$mean_rsm,
                              use = 'complete.obs'),
                          r_n),
             rmse_n = ifelse(w == this_w & temp == this_temp,
                             rmse2(d.means.n$mean_human, d.means.n$mean_rsm),
                             rmse_n),
             r_s = ifelse(w == this_w & temp == this_temp,
                          cor(d.means.s$mean_human, d.means.s$mean_rsm,
                              use = 'complete.obs'),
                          r_s),
             rmse_s = ifelse(w == this_w & temp == this_temp,
                             rmse2(d.means.s$mean_human, d.means.s$mean_rsm),
                             rmse_s)
      )
  }
}

gs.detective %>%
  arrange(w, temp) %>%
  datatable(rownames = F)
```

```{r}
best_w.detective.naive = 0.8
best_temp.detective.naive = 0.05
best_w.detective.sophisticated = 0.5
best_temp.detective.sophisticated = 0.2

data.detective.rsm = import_detective_model_data(best_w.detective.naive,
                                                 best_temp.detective.naive) %>%
  filter(agent_type == 'naive') %>%
  rbind(import_detective_model_data(best_w.detective.sophisticated,
                                    best_temp.detective.sophisticated) %>%
          filter(agent_type == 'sophisticated'))
```

### Uniform simulation
```{r}
data.detective.unif = read_csv(str_c(data_dir, 'exp2_detective/uniform/uniform_model.csv'),
                             show_col_types = F) %>%
  mutate(agent_type = 'naive') %>%
  rbind(read_csv(str_c(data_dir, 'exp2_detective/uniform/uniform_model.csv'),
                 show_col_types = F) %>%
          mutate(agent_type = 'sophisticated')) %>%
  select(-grid, -crumb) %>%
  mutate(id = NA)
```

### Empirical model
```{r}
data.detective.emp = read_csv(str_c(data_dir, 'exp2_detective/',
                                    'empirical/empirical_model.csv'),
                              show_col_types = F) %>%
  select(-grid, -crumb) %>%
  mutate(id = NA)
```

### Empirical mismatched model
```{r}
data.detective.emp_mis = read_csv(str_c(data_dir, 'exp2_detective/',
                                        'empirical_mismatched',
                                        '/empirical_mismatched.csv'),
                                  show_col_types = F) %>%
  select(-grid, -crumb) %>%
  mutate(id = NA)
```

### GPT-4o
```{r}
data.detective.gpt = read_csv(str_c(data_dir, 'exp2_detective/gpt4o/gpt4o.csv'),
                              show_col_types = F) %>%
  select(-ends_with('prompt'))

sum(data.detective.gpt$is_valid) / length(data.detective.gpt$is_valid)

data.detective.gpt = data.detective.gpt %>%
  filter(is_valid) %>%
  select(id, trial, agent_type, response)
```

### Heuristic model
```{r}
data.detective.heu.modeling = read_csv(str_c(data_dir, 'exp2_detective/',
                                             'heuristic/features.csv'),
                                       show_col_types = F)

fit.detective.heu.naive = brm(
  formula = response ~ 1 + fridge_A + crumb_A + fridge_B + crumb_B + (1 | id) + (1 | trial),
  data = data.detective.human %>%
    filter(agent_type == 'naive') %>%
    left_join(data.detective.heu.modeling,
              by = c('trial', 'agent_type')),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.detective.heu.naive'
)

fit.detective.heu.sophisticated = brm(
  formula = response ~ 1 + fridge_A + crumb_A + fridge_B + crumb_B + (1 | id) + (1 | trial),
  data = data.detective.human %>%
    filter(agent_type == 'sophisticated') %>%
    left_join(data.detective.heu.modeling,
              by = c('trial', 'agent_type')),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.detective.heu.sophisticated'
)

data.detective.heu = data.detective.human %>%
  filter(agent_type == 'naive') %>%
  cbind(fit.detective.heu.naive %>%
          fitted(newdata = data.detective.human %>%
                   filter(agent_type == 'naive') %>%
                   left_join(data.detective.heu.modeling,
                             by = c('trial', 'agent_type')),
                 re_formula = NA) %>%
          as_tibble() %>%
          select(pred = Estimate)) %>%
  distinct(trial, pred) %>%
  mutate(agent_type = 'naive') %>%
  rbind(data.detective.human %>%
          filter(agent_type == 'sophisticated') %>%
          cbind(fit.detective.heu.sophisticated %>%
                  fitted(newdata = data.detective.human %>%
                           filter(agent_type == 'sophisticated') %>%
                           left_join(data.detective.heu.modeling,
                                     by = c('trial', 'agent_type')),
                         re_formula = NA) %>%
                  as_tibble() %>%
                  select(pred = Estimate)) %>%
          distinct(trial, pred) %>%
          mutate(agent_type = 'sophisticated')) %>%
  rename(response = pred) %>%
  mutate(id = NA)
```

### Combined
```{r}
data.detective = combine_detective_data(list(
  'human' = data.detective.human,
  'unif' = data.detective.unif,
  'rsm' = data.detective.rsm,
  'emp' = data.detective.emp,
  'emp_mis' = data.detective.emp_mis,
  'gpt4o' = data.detective.gpt,
  'heu' = data.detective.heu
))

data.detective.means = compute_detective_means(data.detective)
```

## Analysis

### Uncertainty
```{r}
fit.detective.human = brm(
  formula = abs_response ~ 1 + agent_type + (1 | id) + (1 | trial),
  data = data.detective %>%
    filter(model == 'human') %>%
    mutate(abs_response = abs(response)),
  iter = 2000,
  seed = 1,
  file = 'cache/fit.detective.human'
)

summary(fit.detective.human)
```

### Model fit comparison
```{r}
# removed one trial because it's NA for GPT 
n_trials = data.detective.means %>% 
  pull(trial) %>% 
  as.character() %>% 
  as.numeric() %>% 
  max() - 1

data.detective.means %>% 
  select(trial, agent_type, contains("mean"))  %>%
  na.omit() %>%
  mutate(residuals_rsm = mean_human - mean_rsm,
         residuals_emp = mean_human - mean_emp,
         residuals_emp_mis = mean_human - mean_emp_mis,
         residuals_gpt4o = mean_human - mean_gpt4o,
         residuals_unif = mean_human - mean_unif,
         residuals_heu = mean_human - mean_heu) %>% 
  select(trial, agent_type, contains("residuals")) %>% 
  pivot_longer(cols = -c(trial, agent_type)) %>% 
  group_by(agent_type, name) %>%
  nest() %>% 
  mutate(log_likelihood = map_dbl(.x = data,
                                  .f = ~ .x %>% 
                                na.omit() %>% 
                                pull(value) %>% 
                                fitdistr(densfun = "normal") %>% 
                                .[["loglik"]])) %>% 
  ungroup() %>%
  mutate(n_parameters = case_when(name == "residuals_rsm" ~ 2,
                                  name == "residuals_heu" ~ 5,
                                  TRUE ~ 0),
         AIC = -2 * log_likelihood + 2 * n_parameters,
         AIC = format(round(AIC, 2), nsmall = 2)) %>%
  select(agent_type, name, AIC) %>% 
  mutate(name = factor(name,
                       levels = str_c("residuals_",
                                      c("rsm", "emp", "emp_mis",
                                        "gpt4o", "unif", "heu")),
                       labels = model_labels[-1])) %>%
  pivot_wider(names_from = agent_type,
              values_from = AIC)
```

## Figures

```{r}
detective.scatterplot = function(data, model, x_title = F, y_title = '', tagg = '') {
  r = cor(data[[str_c('mean_', model)]],
          data[['mean_human']],
          use = 'complete.obs') %>% round(2)
  rmse = rmse2(data[[str_c('mean_', model)]], data[['mean_human']]) %>% round(2)
  
  g = ggplot(data = data,
         aes(x = get(str_c('mean_', model)),
             y = mean_human,
             fill = agent_type)) +
    geom_abline(intercept = 0, slope = 1,
                linetype = 2, linewidth = 0.5) +
    # error bars
    geom_linerange(size = 0.5,
                   mapping = aes(ymin = low_human,
                                 ymax = high_human),
                   color = 'gray50',
                   alpha = 0.4) +
    geom_errorbarh(mapping = aes(xmin = get(str_c('low_', model)),
                                 xmax = get(str_c('high_', model))),
                   color = 'gray50',
                   alpha = 0.4,
                   height = 0) +
    # means
    geom_point(shape = 21,
               size = 2) +
    geom_text(label = sprintf('RMSE = %.2f\nr = %s', rmse, r),
              hjust = 0,   # left align
              x = -50, y = 50,
              size = 4, check_overlap = T) +
    scale_fill_manual(values = agent_type_colors) +
    scale_x_continuous(name = ifelse(x_title, model_labels[model], ''),
                       limits = c(-50, 50)) +
    scale_y_continuous(name = y_title,
                       limits = c(-50, 55),
                       breaks = seq(-50, 50, 25)) +
    coord_cartesian(clip = 'off') +
    labs(tag = tagg) +
    theme(legend.position = 'bottom',
          legend.title = element_blank(),
          legend.text = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          aspect.ratio = 1
          )
  
  return (g)
}
```

```{r fig.width = 13, fig.height = 4.75, warning = F}
scatter.n1 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'rsm',
  y_title = 'Naive humans',
  tagg = 'A'
)

scatter.n2 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'emp',
  tagg = 'B'
)

scatter.n3 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'emp_mis',
  tagg = 'C'
)

scatter.n4 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'gpt4o',
  tagg = 'D'
)

scatter.n5 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'unif',
  tagg = 'E'
)

scatter.n6 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'naive'),
  model = 'heu',
  tagg = 'F'
)

scatter.s1 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'rsm',
  x_title = T,
  y_title = 'Sophisticated humans'
)

scatter.s2 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'emp',
  x_title = T
)

scatter.s3 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'emp_mis',
  x_title = T
)

scatter.s4 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'gpt4o',
  x_title = T
)

scatter.s5 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'unif',
  x_title = T
)

scatter.s6 = detective.scatterplot(
  data.detective.means %>%
    filter(agent_type == 'sophisticated'),
  model = 'heu',
  x_title = T
)

scatter.n1 + scatter.n2 + scatter.n3 + scatter.n4 + scatter.n5 + scatter.n6 + scatter.s1 + scatter.s2 + scatter.s3 + scatter.s4 + scatter.s5 + scatter.s6 +
  plot_layout(guides = 'collect',
              ncol = 6) &
  theme(legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.title = element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.spacing.x = unit(0, 'cm'),
        legend.margin = margin(t = 0, b = 0, unit = 'cm'),
        plot.tag.position = c(0.1, 1.05),
        plot.margin = margin(t = 0.1, r = 0.1, b = -0.1, l = 0.1, unit = 'cm')
        )

# ggsave(str_c(figures_dir, 'detective.pdf'), width = 13, height = 4.75)
```

# Session info
```{r}
sessionInfo()
```


